Draft of paper

### 2024-12-12 Paper draft/draft.tex
Optimization sector finished

\begin{figure*}
\begin{subequations}\label{eqn:LMI}
\begin{align}\label{eqn:lyapunov}
  &\Xi =\bmx{
  \begin{array}{c|c}
  \bmx{\bar{A}\tr\\
  \bar{B}\tr\\
  C\tr
  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
  P & 0 & 0\\
  0 & 0 & 0\\
  0 & 0 & 0
  } - (\1_l \otimes R_{\nu})\tr \He{\bX} (\1_l \otimes R_{\nu})
   + R_s\tr \bmx{
    S & Q\\
    \star & W
  }  R_s\hspace{.1em}& 0\\[1.5em]
\hline \\[-0.9em]
  0 & \hspace{.1em} 2 (\bR - \Id)
  \end{array}} < 0\\
  &\bmx{
    P & {Z_j^i}\tr\\
    \star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
  } \geq 0 \ \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
  &{R^i}\tr\! \He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
  &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
\end{align}
\end{subequations}
\vspace{.5em}
\hrule
\end{figure*}
% \begin{figure*}
% \begin{equation}\label{eqn:first-step}
%  \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
%   \bmx{\bar{A}\tr\\
%   \bar{B}\tr\\
%   C\tr
%   } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
%   P & 0 & 0\\
%   0 & 0 & 0\\
%   0 & 0 & 0
%   } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}
%  - \pmx{\1_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi}
%  + \bmx{\tilde{x}\\ \tilde{\Phi}}\tr\! \bmx{S & Q\\ \star & W} \bmx{\tilde{x} \\ \tilde{\Phi}}
%  + \sqrt{\boldsymbol{\eta}}\tr \He{\bR - \Id} \sqrt{\boldsymbol{\eta}}
% \end{equation}
% \vspace{.5em}
% \hrule
% \end{figure*}

@@

We can define the following definite-positive candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + \He{\1_l\tr\!\boldsymbol{\eta}}$. 
By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)}\tr = R_s \! \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)}\tr$ then the following holds $\forall x \in \cE(P, x_*)$:
\begin{equation}\label{eqn:sector-sign}
\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)}\tr \! R_s\tr \! \bmx{S & Q \\ \star & W} R_s \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)} \geq 0
\end{equation}
With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting \\$\xi = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr = R_{\nu} \zeta$, if we pre and post multiply \eqref{eqn:lyapunov} by $\bmx{\zeta\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose, by expressing the midterm as follows: 
\begin{equation}
\begin{aligned}
  &\zeta\tr \pmx{\1_l \otimes R_{\nu}}\tr \He{\bX} \pmx{\1_l \otimes R_{\nu}} \zeta = \\
  =& \pmx{\1_l 1 \otimes R_{\nu} \zeta}\tr \He{\bX} \pmx{\1_l 1 \otimes R_{\nu} \zeta} = \\
  =& \pmx{\1_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi} = \\
  =& \1_l\tr \pmx{\Id_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi} = \1_l\tr \mathbf{\Psi_x}
\end{aligned}
\end{equation}
and considering \eqref{eqn:sector-sign} we verify Assumption \ref{ass:nonlin} then by collecting the proper terms $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$ we have
\begin{equation}
\begin{aligned}
&\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
  \bmx{\bar{A}\tr\\
  \bar{B}\tr\\
  C\tr
  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
  P & 0 & 0\\
  0 & 0 & 0\\
  0 & 0 & 0
  } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}} + \\
  &+ \He{\1_l\tr \bmx{(\bR - \Id) \boldsymbol{\eta} - \mathbf{\Psi_x}}} = \\
  &= (\tilde{x}^+)^{\top} P \tilde{x}^+ - \tilde{x}\tr P \tilde{x} + \He{\1_l\tr (\boldsymbol{\eta}^+ - \boldsymbol{\eta})} = \Delta V < 0, 
\end{aligned}
\end{equation}
{\color{blue} Therefore, $\cE(P, x_*)$ is forward invariant and all trajectories starting inside such a set converge to $(x, \0)$, thus concluding the proof.}
\end{proof}
\subsection{Optimization procedure}

>>>>

\begin{figure*}
\begin{subequations}\label{eqn:LMI}
\begin{align}\label{eqn:lyapunov}
  &\Xi =\bmx{
  \begin{array}{c|c}
  \bmx{\bar{A}\tr\\
  \bar{B}\tr\\
  C\tr
  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
  P & 0 & 0\\
  0 & 0 & 0\\
  0 & 0 & 0
  } - (\1_l \otimes R_{\nu})\tr \He{\bX} (\1_l \otimes R_{\nu})
   + R_s\tr \bmx{
    S & Q\\
    \star & W
  }  R_s\hspace{.1em}& 0\\[1.5em]
\hline \\[-0.9em]
  0 & \hspace{.1em} 2 (\bR - \Id)
  \end{array}} < 0\\
  &\bmx{
    P & {Z_j^i}\tr\\
    \star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
  } \geq 0 \qquad \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
  &{R^i}\tr\! \He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \qquad \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
  &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \qquad \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
\end{align}
\end{subequations}
\vspace{.5em}
\hrule
\end{figure*}
% \begin{figure*}
% \begin{equation}\label{eqn:first-step}
%  \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
%   \bmx{\bar{A}\tr\\
%   \bar{B}\tr\\
%   C\tr
%   } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
%   P & 0 & 0\\
%   0 & 0 & 0\\
%   0 & 0 & 0
%   } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}
%  - \pmx{\1_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi}
%  + \bmx{\tilde{x}\\ \tilde{\Phi}}\tr\! \bmx{S & Q\\ \star & W} \bmx{\tilde{x} \\ \tilde{\Phi}}
%  + \sqrt{\boldsymbol{\eta}}\tr \He{\bR - \Id} \sqrt{\boldsymbol{\eta}}
% \end{equation}
% \vspace{.5em}
% \hrule
% \end{figure*}

@@

We can define the following definite-positive candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + \He{\1_l\tr\!\boldsymbol{\eta}}$. 
By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)}\tr = R_s \! \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)}\tr$ then the following holds $\forall x \in \cE(P, x_*)$:
\begin{equation}\label{eqn:sector-sign}
\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)}\tr \! R_s\tr \! \bmx{S & Q \\ \star & W} R_s \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)} \geq 0
\end{equation}
With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting \\$\xi = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr = R_{\nu} \zeta$, if we pre and post multiply \eqref{eqn:lyapunov} by $\bmx{\zeta\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose, by expressing the midterm as follows: 
\begin{equation}
\begin{aligned}
  &\zeta\tr \pmx{\1_l \otimes R_{\nu}}\tr \He{\bX} \pmx{\1_l \otimes R_{\nu}} \zeta = \\
  =& \pmx{\1_l 1 \otimes R_{\nu} \zeta}\tr \He{\bX} \pmx{\1_l 1 \otimes R_{\nu} \zeta} = \\
  =& \pmx{\1_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi} = \\
  =& \1_l\tr \pmx{\Id_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi} = \1_l\tr \mathbf{\Psi_x}
\end{aligned}
\end{equation}
and considering \eqref{eqn:sector-sign} we verify Assumption \ref{ass:nonlin} then by collecting the proper terms $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$ we have
\begin{equation}
\begin{aligned}
&\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
  \bmx{\bar{A}\tr\\
  \bar{B}\tr\\
  C\tr
  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
  P & 0 & 0\\
  0 & 0 & 0\\
  0 & 0 & 0
  } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}} + \\
  &+ \He{\1_l\tr \bmx{(\bR - \Id) \boldsymbol{\eta} - \mathbf{\Psi_x}}} = \\
  &= (\tilde{x}^+)^{\top} P \tilde{x}^+ - \tilde{x}\tr P \tilde{x} + \He{\1_l\tr (\boldsymbol{\eta}^+ - \boldsymbol{\eta})} = \Delta V < 0, 
\end{aligned}
\end{equation}
{\color{blue} Therefore, $\cE(P, x_*)$ is forward invariant and all trajectories starting inside such a set converge to $(x, \0)$, thus concluding the proof.}
\end{proof}

\subsection{Optimization procedure}
We state now a corollary whose results will be useful to derive some performances upgrade that will be later discussed.

\begin{corollary}\label{cor:optimization}
  \emph{Given Theorem \ref{thm:stability}, considering the decision variables $\Sigma, P, T, Z, \bX, \bR, S, Q, W, N^1_i, N^2_i, N^3_i, \beta_i \quad \forall i \in \left\{ 1, \dots, l \right\}$ with $\Sigma$ being a diagonal matrix $\Sigma > 0 \in \RR^{(n_x + 2 n_{\phi} + n_q) \times (n_x + 2 n_{\phi} + n_q)}$, and $\beta_i$ being scalars with the following optimization scheme:}
  \begin{equation}\label{eqn:optimization}
  \begin{aligned}
    &\min \trace{P} + \trace{\Sigma} + \sum_{i=1}^{l} \beta_i\\
    &\text{subject to } \eqref{eqn:LMI},\\ 
    &\qquad \qquad \Xi + \Sigma \geq 0,\\
    &\qquad \qquad \bmx{-\beta_i \Id & X^i\\
    {X^i}\tr & -\Id} \leq 0 \qquad i \in \left\{ 1, \dots, l \right\}
  \end{aligned}
  \end{equation}
  \emph{Then $(x_*, \0)$ is LES and $\cE(P, x_*)$ is an inner approximation of the ROA.}
\end{corollary}

By adding these constraints we obtain some desirable performances upgrade:
\begin{itemize}
  \item By minimizing the trace of $P$ we are maximizing the volume of the ellipsoid $\cE(P, x_*)$ hence the ROA.
  \item By minimizing the trace of $\Sigma$ we are minimizing the conservatism of the controller by obtaining a constraint matrix $\Xi$ directly related to the decay rate of the Lyapunov function as close to zero as possible but never greater due to the second condition in \eqref{eqn:optimization}.
  \item By taking \eqref{eqn:optimization} and applying the Schur complement to the last condition we obtain
  $$
  X^i {X^i}\tr \leq \beta_i \Id
  $$
  meaning that the smaller the value of $\beta_i$ the smaller will be the value that appears directly in the ETM triggering condition \eqref{eqn:etm-trigger} hence the lower the number of events.
\end{itemize}

Let us comment the quantity $\alpha$ that appears in \eqref{eqn:inclusion}, if this was considered as a variable then the statements of Theorem \ref{thm:stability} would not hold since we would encounter some bilinear terms. By fixing $\alpha$ to a constant value before the LMI evaluation we address the problem. The choice of $\alpha$ value is crucial since it will directly affect the conservatism of the solution that the problem will provide. The followed approach sees the deployment of a \emph{golden ratio search} to extract a good value. The algorithm is based on the bisection method with some basic improvements:
\begin{itemize}
  \item Using the golden ratio to divide the interval reduces the number of LMI solutions compared to the bisection method. Converging faster since it narrows the search interval more efficiently.
  \item It maintains a constant reduction factor for the interval size, leading to more consistent progress towards the optimal value.
  \item It is particularly useful for unimodal functions like the one we are dealing with after we set the volume of the ROA as the parameter to optimize.
\end{itemize}
